Effectful Semantics for Substructural SSA

### 2023-11-14 pldi24/paper.tex
Tap

\TODO{intro text}

\begin{lemma}[Weakening] \
  \begin{itemize}
    \item If \(\cwk{\Gamma}{\Delta}\) and \(q \subseteq r\),
    \(\hasty{\Delta}{p}{a}{A}{r}\), then \(\hasty{\Gamma}{p}{a}{A}{q}\)
    \item If \(\cwk{\Gamma}{\Delta}\), \(\haslb{\Delta}{t}{\ms{L}}\), then \(\haslb{\Gamma}{t}{\ms{L}}\)
    \item If \(\lwk{\ms{L}}{\ms{K}}\), \(\haslb{\Gamma}{t}{\ms{L}}\) then \(\haslb{\Gamma}{t}{\ms{K}}\)
  \end{itemize}
\end{lemma}

@@

>>>>

We can now state some of the basic metatheoretic properties of \isotopessa. We
begin with weakening, which we state in our setting as follows:
\begin{lemma}[Weakening] \
  \begin{itemize}
    \item If \(\cwk{\Gamma}{\Delta}\) and \(q \subseteq r\),
    \(\hasty{\Delta}{p}{a}{A}{r}\), then \(\hasty{\Gamma}{p}{a}{A}{q}\)
    \item If \(\cwk{\Gamma}{\Delta}\), \(\haslb{\Delta}{t}{\ms{L}}\), then \(\haslb{\Gamma}{t}{\ms{L}}\)
    \item If \(\lwk{\ms{L}}{\ms{K}}\), then:
    \begin{itemize}
      \item If \(\haslb{\Gamma}{t}{\ms{L}}\) then \(\haslb{\Gamma}{t}{\ms{K}}\)
      \item If \(\lhaslb{\ms{W}}{L}{\ms{L}}\) then \(\lhaslb{\ms{W}}{L}{\ms{L}}\)
    \end{itemize}
  \end{itemize}
\end{lemma}
That is, we can do the following and still keep our typing judgements valid
\begin{itemize}
  \item Add unused \textit{affine} variables
  \item \textit{Reduce} the quantity of an expressions type
  \item Add arbitrary unused labels, and remove unused \textit{affine} variables
  from existing labels
\end{itemize}
This can be proved by a relatively straightforward mutual induction.

Our main metatheoretic results will be \textit{substitution} and
\textit{label-substitution}, which will require introducing some new
metatheoretic judgements to state properly, which we do in Figure
\ref{fig:ssa-meta-judgements}.

@@

\begin{lemma}[Substitution] 
  Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
  \begin{itemize}
    \item If \(\hasty{\Gamma}{p}{a}{A}{q}\), then \(\hasty{\Theta}{p}{[\gamma]a}{A}{q}\) and \(\hasty{\Theta}{p}{\exprletsubst{\substctx{\gamma}{\Gamma}}{a}}{A}{q}\)
    \item If \(\haslb{\Gamma}{t}{\ms{L}}\), then there exists \(\ms{K}\) such
    that \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) and
    \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\), \(\haslb{\Theta}{\stmtletsubst{\substctx{\gamma}{\Gamma}}{t}}{\ms{K}}\)
    \item If \(\lhaslb{\ms{W}}{L}{\ms{L}}\), then there exists \(\ms{K},
    \ms{W}'\) such that 
    \begin{itemize}
      \item \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) 
      \item \(\lbsubst{\substlbs{\gamma}{\ms{W}'}}{\ms{W}}{\ms{W}'}\) 
      \item \(\lhaslb{\ms{W}'}{[\gamma]L}{\ms{K}}\)
    \end{itemize}
  \end{itemize}
\end{lemma}
